---
title: "Pilot Preliminary Analysis"
author: "John Bocharov, Sho Ohota, Apekshit Sharma"
date: "April 11, 2016"
output: pdf_document
---

# Prerequisites

Load the required libraries

```{r, message=FALSE}
library(lubridate)
library(plyr)
library(reshape2)
```

Load the data:

```{r}
setwd('~/datasci/W241/w241-political-news/')
raw_data <- read.csv("241_Project__Pilot_mTurk.csv", na.strings='')
```


# Data Cleaning

The most interesting decision is which observations to keep. We keep only people who took > 100 seconds to complete the survey, as an indicator of actual well-considered responses.

```{r}

perform_data_cleaning <- function (raw_data) {
  
  # Not used in at runtime, but useful for debugging this function
  questions <- raw_data[1,]
  
  # Remove the question text row
  working_data <- raw_data[ 2:nrow(raw_data), ]
  
  start_time <- ymd_hms(working_data$V8)
  completion_time <- ymd_hms(working_data$V9)
  working_data$time_taken <- as.numeric( completion_time - start_time )
  
  working_data <- working_data[ working_data$time_taken > 100, ]
  
  working_data <- rename(working_data, c(
    "Q18_4" = "agreement_fox_correctly_labeled",
    "Q20_7" = "agreement_fox_counter_labeled",
    "Q18_5" = "credibility_fox_correctly_labeled",
    "Q20_9" = "credibility_fox_counter_labeled",
    "Q19_4" = "agreement_huff_correctly_labeled",
    "Q18_6" = "agreement_huff_counter_labeled",
    "Q19_5" = "credibility_huff_correctly_labeled",
    "Q18_7" = "credibility_huff_counter_labeled"
  ))
  
  likert <- levels(working_data$agreement_fox_correctly_labeled)
  likert_order <- c("Strongly Disagree", "Disagree", "Neither Agree nor Disagree", "Agree", "Strongly Agree")
  to_likert_factor <- function(ratings) {
    return(
      factor(likert[ratings], levels=likert_order)
    )
  }
  
  merge_responses <- function (correctly_labeled, counter_labeled) {
    return(ifelse(
      is.na(counter_labeled), correctly_labeled, counter_labeled
    ))
  }
  
  agreement_fox_raw <- merge_responses( 
    working_data$agreement_fox_correctly_labeled,
    working_data$agreement_fox_counter_labeled
  )
  
  credibility_fox_raw <- merge_responses(
    working_data$credibility_fox_correctly_labeled,
    working_data$credibility_fox_counter_labeled
  )
  
  agreement_huff_raw <- merge_responses(
    working_data$agreement_huff_correctly_labeled,
    working_data$agreement_huff_counter_labeled
  )
  
  credibility_huff_raw <- merge_responses(
    working_data$credibility_huff_correctly_labeled,
    working_data$credibility_huff_counter_labeled
  )
  
  working_data$agreement_fox <- to_likert_factor(agreement_fox_raw)
  working_data$credibility_fox <- to_likert_factor(credibility_fox_raw)
  working_data$agreement_huff <- to_likert_factor(agreement_huff_raw)
  working_data$credibility_huff <- to_likert_factor(credibility_huff_raw)
  
  working_data <- rename(working_data, c(
    "Q1" = "party_dem_ind_rep",
    "Q2" = "raw_party_loyalty",
    "Q2.1" = "party_dem_rep"
  ))
  
  party <- working_data$party_dem_ind_rep
  
  ##Assign Independents which party they feel more affiliated to
  party[party=="Independent"] <- working_data$party_dem_rep[
    !is.na(working_data$party_dem_rep)
  ] 
  
  working_data$party <- factor(party)
  
  party_loyalty <- working_data$raw_party_loyalty
  
  party_loyalty[is.na(party_loyalty)] <- "Weak"
  working_data$party_loyalty <- factor(party_loyalty, levels=c("Weak", "Moderate", "Strong"))
  
  working_data$treatment <- ifelse(is.na(working_data$agreement_fox_counter_labeled), 0, 1)
  
  return(
    working_data[,c(
      "mTurkCode", 
      "treatment", 
      "agreement_huff", 
      "credibility_huff", 
      "agreement_fox", 
      "credibility_fox",
      "party",
      "party_loyalty"
    )]
  )
}

clean_data <- perform_data_cleaning(raw_data)
```

Show a small sample of the resulting data:

```{r}
head(clean_data)
```

Some key model specifications (courtesy of Sho):

```{r}
summary(lm(as.numeric(agreement_fox) ~ party + party_loyalty + treatment, data=clean_data))

summary(lm(as.numeric(credibility_fox) ~ party + party_loyalty + treatment, data=clean_data))

summary(lm(as.numeric(agreement_huff) ~ party + party_loyalty + treatment, data=clean_data))

summary(lm(as.numeric(credibility_huff) ~ party + party_loyalty + treatment, data=clean_data))
```

# AP Score imputation

First we'll need a utility function to calculate the label from a source vector and a treatment assignment vector.

```{r}
compute_label <- function(source, treatment) {
  counter_source <- ifelse(source == "Huff", "Fox", "Huff")
  return(factor(
    ifelse(
      treatment == 1, 
      counter_source, 
      as.character(source))
  ))
}
```

Since we did not ask the AP question directly in the pilot, impute AP as halfway between the Fox and HuffPo scores. 

Then we measure the difference between the score of each source (Fox, HuffPo) against the imputed AP rating as the lift for that source.

Let's look at the imputed agreement ratings.

```{r}
impute_agreement <- function(clean_data) {
  mTurkCode <- clean_data$mTurkCode
  treatment <- clean_data$treatment
  agreement_fox <- as.numeric(clean_data$agreement_fox)
  agreement_huff <- as.numeric(clean_data$agreement_huff)
  imputed_agreement_ap <- ( agreement_fox + agreement_huff ) / 2.0
  
  party <- clean_data$party
  party_loyalty <- clean_data$party_loyalty
  
  imputed_agreement <- data.frame( 
    mTurkCode = mTurkCode, 
    treatment = treatment,
    party = party,
    party_loyalty = party_loyalty,
    imputed_ap = imputed_agreement_ap,
    Fox = agreement_fox - imputed_agreement_ap,
    Huff = agreement_huff - imputed_agreement_ap
  )
  
  return(melt( 
    imputed_agreement,
    id.vars = c("mTurkCode", "treatment", "party", "party_loyalty", "imputed_ap"),
    measure.vars = c("Fox", "Huff"),
    variable.name = "source",
    value.name = "lift"
  ))
}

imputed_agreement <- impute_agreement(clean_data)
imputed_agreement$label <- compute_label(imputed_agreement$source, imputed_agreement$treatment)

head(imputed_agreement)
```

## Agreement Model specifications

The first interesting spec is the lift by source and party

```{r}
summary(lm( lift ~ source + party + source:party, data = imputed_agreement ))
```

Strange conclusions:

(a) HuffPo is significantly more agreed with, and
(b) especially by Republicans (very counter-intuitive)

Then we add the label to the regression

```{r}
summary(lm( lift ~ source + party + source:party + label, data = imputed_agreement ))
```

With a p-value of 0.40, nowhere near significance. Then we add the label:source interaction:

```{r}
summary(lm( lift ~ source + party + source:party + label + label:source, data = imputed_agreement ))
```

And finally the label:source:party interaction:

```{r}
summary(lm( lift ~ source + party + source:party + label + label:source:party, data = imputed_agreement ))
```



## Imputed Credibility

Then we also impute the credibility:

```{r}
impute_credibility <- function(clean_data) {
  mTurkCode <- clean_data$mTurkCode
  treatment <- clean_data$treatment
  credibility_fox <- as.numeric(clean_data$credibility_fox)
  credibility_huff <- as.numeric(clean_data$credibility_huff)
  imputed_credibility_ap <- ( credibility_fox + credibility_huff ) / 2.0
  
  party <- clean_data$party
  party_loyalty <- clean_data$party_loyalty
  
  imputed_credibility <- data.frame( 
    mTurkCode = mTurkCode, 
    treatment = treatment,
    party = party,
    party_loyalty = party_loyalty,
    imputed_ap = imputed_credibility_ap,
    Fox = credibility_fox - imputed_credibility_ap,
    Huff = credibility_huff - imputed_credibility_ap
  )
  
  return(melt( 
    imputed_credibility,
    id.vars = c("mTurkCode", "treatment", "party", "party_loyalty", "imputed_ap"),
    measure.vars = c("Fox", "Huff"),
    variable.name = "source",
    value.name = "lift"
  ))
}

imputed_credibility <- impute_credibility(clean_data)
imputed_credibility$label <- compute_label(imputed_credibility$source, imputed_credibility$treatment)

head(imputed_credibility)
```

First spec, lift by source and party:

```{r}
summary(lm( lift ~ source + party + source:party, data = imputed_credibility ))
```

Again, strange conclusions:

(a) Republicans more likely to give credibility to the source
(b) especially by HuffPo (very counter-intuitive)

Then we add the label to the regression

```{r}
summary(lm( lift ~ source + party + source:party + label, data = imputed_credibility ))
```

With a p-value of 0.23, still not significant. Then we add the label:source interaction:

```{r}
summary(lm( lift ~ source + party + source:party + label + label:source, data = imputed_credibility ))
```

And finally the label:source:party interaction:

```{r}
summary(lm( lift ~ source + party + source:party + label + label:source:party, data = imputed_credibility ))
```